02_DESeq2

### 2020-02-18 analysis/02_DESeq2.Rmd
prl degs

```{r readfiles, eval=F, include=F}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

@@

```{r manipulation1, include = F}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

@@

plotDispEsts( dds, ylim = c(1e-6, 1e1) )

@@

restoDEGs <- function(res, up, down){
  DEGs <- data.frame(gene = row.names(res),
                        padj = res$padj, 
                        logpadj = -log10(res$padj),
                        lfc = res$log2FoldChange)
  DEGs <- na.omit(DEGs)
  DEGs <- DEGs %>%
    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
                                                           yes = down, no = "NS"))) %>% 
    dplyr::arrange(gene)
  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up)) 
  # write DEGsframe of only significant genes
  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
  print(str(DEGs))
  #partialfilename = paste("_", down, "_", up, sep = "")
  #myfilename = paste0("../results/DESeq2/", mytissue, partialfilename, "_DEGs.csv")
  #write.csv(DEGs, myfilename, row.names = F)
  print(head(DEGs))
  return(DEGs)
}

@@

top <- plotbarvolcano(sexdegs, c("a","b"), "sex")
bottom <- plotbarvolcano(PRLdegs, c("c","d"), "prolactin expression")

plot_grid(top,bottom, nrow = 2)

head(PRLdegs)
head(sexdegs)

@@

ggplot(sexPRLdesg, aes(x = direction.x, fill = direction.y)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = allcolors, na.value= "#bdbdbd",
                    labels= c("low PRL", "hi PRL", "NS"),
                    name = "PRL DEGs") +
  scale_x_discrete(breaks = c("female", "male", NA),
                   labels = c("female", "male", "NS"),
                   name = "Sex DEGs")

interaction <- sexPRLdesg %>% filter(direction.x %in% c("male", "female") & 
                                       direction.y %in% c("hi", "lo"))
interaction

sexonly <- sexPRLdesg %>% filter(direction.x %in% c("male", "female") & 
                                       !direction.y %in% c("hi", "lo") )
sexonly
PRLonly <- sexPRLdesg %>% filter(!direction.x %in% c("male", "female")& 
                                       direction.y %in% c("hi", "lo"))
PRLonly

write.csv(sexPRLdesg, "../results/DEG-PRLsex.csv")
```

>>>>

```{r readfiles, eval=T}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

@@

```{r manipulation1, eval = F, include = F}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

@@

@@

restoDEGs <- function(res, up, down){
  DEGs <- data.frame(gene = row.names(res),
                        padj = res$padj, 
                        logpadj = -log10(res$padj),
                        lfc = res$log2FoldChange)
  DEGs <- na.omit(DEGs)
  DEGs <- DEGs %>%
    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
                                                           yes = down, no = "NS"))) %>% 
    dplyr::arrange(gene)
  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up)) 
  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
  print(head(DEGs))
  return(DEGs)
}

@@

top <- plotbarvolcano(sexdegs, c("a","b"), "sex")
bottom <- plotbarvolcano(PRLdegs, c("c","d"), "PRL")

ab <- plot_grid(top,bottom, nrow = 2)

sexPRLdesg <- full_join(sexdegs, PRLdegs, by = "gene")

@@

write.csv(sexPRLdesg, "../results/DEG-PRLsex.csv")

sexPRLdesg2 <- sexPRLdesg %>%
  mutate(higherin = paste(direction.x, direction.y, sep = "_")) %>%
  mutate(direction = paste(higherin, "PRL", sep = "")) %>%
  select(gene, direction) %>%
  group_by(direction) %>%
  summarize(DEGs = str_c(gene , collapse = " "))
sexPRLdesg2
write.csv(sexPRLdesg2, "../results/PRLsexinteraction.csv")

c <- ggplot(sexPRLdesg, aes(x = direction.x, fill = direction.y)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = allcolors, na.value= "#bdbdbd",
                    labels= c("lo", "hi", "NS"),
                    name = "PRL") +
  scale_x_discrete(breaks = c("female", "male", NA),
                   labels = c("female", "male", "NS"),
                   name = "Sex")  +
  labs(y = "Total DEGs") +
  theme_B3() +
  theme(legend.position = "top")

plot_grid(ab, c, nrow = 1, rel_widths = c(2,1.2), labels = c(" ", "d"), label_size = 12)
